---
title: "Assignment of week 2"
author: "Pieter ten Have"
date: "October 24, 2016"
output: html_document
---

This document describes the analysis of the activity dataset.

```{r setup, echo=TRUE}
rm(list=ls())
setwd("C:/Users/Pieter/Documents")
a<-read.csv(file="activity.csv", header=T, na.string="NA")
steps_a_day<-tapply(a$steps, a$date, sum, na.rm=T)
steps_a_day<-as.numeric(steps_a_day)
hist(steps_a_day, xlab="Number of steps per day")
```

Calculate the median.

```{r}
median(steps_a_day, na.rm=T)

```

Calculate the average.

```{r}
mean(steps_a_day, na.rm=T)
```

Below you see a plot of the average number of steps in each 5-min interval of the day.

```{r}
a$interval<-as.numeric(a$interval)
steps_per_interv<-tapply(a$steps, a$interval, mean, na.rm=T)
interv<-unique(a$interval)
steps_tbl<-data.frame(interv, steps_per_interv)
plot(steps_tbl, type="n", xlab="5 min intervals from 0 AM to 12 PM",
     main="Average number of steps in 5 min interval during the day")
lines(steps_tbl)
```

The code below imputes missing values. It looks for the average number of steps in a corresponding interval in the other days, and replaces the missing value by this number.

```{r}
for (lst in interv){
  assgn<-steps_tbl$steps_per_interv[steps_tbl$interv==lst]
  a$steps[is.na(a$steps) & a$interval==lst]<-assgn
}
```

Below is a histogram of the average number of steps per day, after imputing the missing values.

```{r}
steps_a_day2<-tapply(a$steps, a$date, sum, na.rm=T)
steps_a_day2<-as.numeric(steps_a_day2)
hist(steps_a_day2, main="Histogram of total number of steps per day")
```

Below is a plot of a weekend day and a non-weekend day with the  average number of steps in each 5-min interval of the day.

```{r}
a$weekday<-weekdays(as.Date(a$date))
a$weekend<-0
a$weekend[a$weekday %in% c("Saturday","Sunday")]<-1

par(mfrow=c(1,2)) 

a10<-subset(a, weekend==0)
steps_per_interv<-tapply(a10$steps, a10$interval, mean)
interv<-unique(a$interval)
steps_tbl<-data.frame(interv, steps_per_interv)
plot(steps_tbl, type="n", main="No. of steps per 5-min
  interval in non-weekend day")
lines(steps_tbl)

a11<-subset(a, weekend==1)
steps_per_interv2<-tapply(a11$steps, a11$interval, mean)
interv<-unique(a$interval)
steps_tbl2<-data.frame(interv, steps_per_interv2)
plot(steps_tbl2, type="n", main="No. of steps per 5-min
interval in weekend day")
lines(steps_tbl2)
```




